### 2021-09-29 documentos/math.tex
For the statistics only

\section{Mathematical derivation} % {{{
\subsection{Diagonalization of Choi matrix} % {{{

Prueba de la $\sa$ \todo{Alejo escribe esta parte. Dejelo aca todo y luego si acaso 
pasamos cosas al apendice. Caso general}

@@

\subsubsection{Single qubit Pauli channels}

\afnote{Notación para matrices densidad vectorizadas? $\varrho$ ?}
\janote{Qué tal $\vec\rho\quad$?}
\afnote{Por ahora uso este $\ket{\rho}$}

We now turn to find the matrix form of the map $\mathcal{E}$ in order to construct the Choi matrix and the conditions the set of parameters $\{\tau_{\alpha}\}$ need to satisfy in order to $\mathcal{E}$ be considered a quantum channel. By writing the density matrix in vector form $\ket{\rho}=2^{-1}\sum_{\alpha=0}^{3} r_{\alpha} \ket{\sigma_{\alpha}}$, the matrix form of the map $\mathcal{E}$ may be written as
% 
\begin{equation}
\hat{\mathcal{E}} = \frac{1}{2}\sum_{\alpha=0}^{3} \tau_{\alpha} \dyad{\sigma_{\alpha}},
% \label{Map_Pauli}
\end{equation}
%
where the kets $\ket{\sigma_{\alpha}}$ correspond to the vectorized elements of the Pauli basis and satisfy the usual relation $ \braket{{\sigma_{\alpha}}}{{\sigma_{\alpha'}}} = \tr \left(\sigma_{\alpha}^{\dagger} \sigma_{\alpha'}\right) = 2 ~\delta_{\alpha\alpha'}$. After some steps it is possible to show that the associated Choi matrix reads
% 
\begin{equation}
 \mathcal{D} = \frac{1}{2}\sum_{\alpha=0}^{3} \tau_{\alpha} \sigma_{\alpha} \otimes \sigma_{\alpha}^*.
%  \label{Choi_1}
\end{equation}
% 
Moreover, observe that $\mathcal{D}$ has the same form as the matrix representation of a random unitary channel, with Kraus operators given by $K_{\alpha}=\sqrt{\tau_{\alpha}/2}\sigma_{\alpha}$ \cite{CHRUSCINSKI20131425}. In this way the Choi matrix $\mathcal{D}$ is diagonal in the Pauli basis (see Appendix for details),
%
\begin{equation}
 \mathcal{D} \ket{\sigma_{\alpha}}= \lambda_{\alpha} \ket{\sigma_{\alpha}},
\end{equation}
% 
with eigenvalues given by 
% 
\begin{equation}
 \lambda_{\alpha} = \sum_{\beta=0}^3\sa_{\alpha\beta}\tau_{\beta},
\end{equation}
%
or more compactly $\vec{\lambda} = \sa~\vec{\tau}$, where $\sa=2^{-1}H\otimes H$ and $H$ is the Hadamard matrix,
%
\begin{equation}
 H=\begin{pmatrix}
        1 & 1\\
        1 & -1\\
    \end{pmatrix}.
 \label{Hadd_Mat}
\end{equation}
% 
\afnote{Podemos definir la matriz de Hadamard con un factor $2^{-1/2}$, o sea $H'=H/\sqrt{2}$ para evitar cargar factores extra hasta el final, de esta manera tenemos simplemente $\san=H^{\otimes 2N}$. Qué opinan?}

\subsubsection{$N$ qubits}
% 
Analogously to the single qubit case, one can define the $N$-qubits Pauli map $\mathcal{E}_N$
% 
\begin{equation}
\rho \to \rho'=\mathcal{E}_N[\rho ]=\frac{1}{2^N}\sum_{\vec{\alpha}} \tau_{\vec{\alpha}} r_{\vec{\alpha}} \sigma_{\vec{\alpha}},
% \label{Pauli_Map}
\end{equation}
% 
with $-1\leq \tau_{\vec{\alpha}} \leq 1$. 
% 
In this case the density matrix in vector form holds 
% 
\begin{equation}
 \ket{\rho}=\frac{1}{2^N}\sum_{\vec{\alpha}} r_{\vec{\alpha}} \ket{\sigma_{\vec{\alpha}}},
\end{equation}
% 
where $\ket{\sigma_{\vec{\alpha}}}=\ket{\sigma_{\alpha_1}}\otimes\dots \otimes\ket{\sigma_{\alpha_N}}$, and satisfy the usual orthogonality relation $\braket{{\sigma_{\vec{\alpha}}}}{{\sigma_{\vec{\alpha}'}}} = \tr \left(\sigma_{\vec{\alpha}}^{\dagger} \sigma_{\vec{\alpha}'}\right) = 2^N ~\delta_{\vec{\alpha}\vec{\alpha}'}$. The matrix representation of this map is given by
% 
\begin{equation}
\hat{\mathcal{E}}_N = \frac{1}{2^N}\sum_{\vec{\alpha}} \tau_{\vec{\alpha}} \dyad{\sigma_{\vec{\alpha}}}.
% \label{Map_Pauli}
\end{equation}
%
As in the previous case, the Choi matrix $\mathcal{D}_N$ may be written in terms of tensor products of Pauli matrices
% 
\begin{equation}
 \mathcal{D}_N = \frac{1}{2^N}\sum_{\vec{\alpha}} \tau_{\vec{\alpha}}  \bigotimes_{j=1}^N \sigma_{\alpha_j} \otimes \sigma_{\alpha_j}^*.
%  \label{Choi_1}
\end{equation}
% 
After several steps, it is possible to prove that analogously to the single qubit case, $\mathcal{D}_N$ is diagonal in the $N$-qubits Pauli basis
%
\begin{equation}
 \mathcal{D}_N \ket{\sigma_{\vec{\alpha}}}= \lambda_{\vec{\alpha}} \ket{\sigma_{\vec{\alpha}}},
\end{equation}
% 
with eigenvalues $\lambda_{\vec{\alpha}}$ given by 
% 
\begin{equation}
 \lambda_{\vec{\alpha}} = \sum_{\vec{\beta}} \san_{\vec{\alpha}\vec{\beta}}\tau_{\vec{\beta}},
\end{equation}
%
where 
% 
\begin{equation}
 \san=\sa^{\otimes N}=\frac{1}{2^N}\left(H\otimes H\right)^{\otimes N}=\left(\frac{H}{\sqrt{2}}\right)^{\otimes 2N}.
\end{equation}
%

(Proofs in an appendix using binary indices XD)

\subsubsection{Appendix - Diagonalization of Choi Matrix $\mathcal{D}_N$}

% }}}

\input{vector_space}

>>>>

\section{Mathematical considerations} 
\subsection{Diagonalization of Choi matrix}

% Intro {{{
Prueba de la $\sa$ \todo{Alejo escribe esta parte. Dejelo aca todo y luego si acaso 
pasamos cosas al apendice. Caso general}

@@

% }}}
\subsubsection{Single qubit Pauli channels} % {{{

\afnote{Notación para matrices densidad vectorizadas? $\varrho$ ?}
\janote{Qué tal $\vec\rho\quad$?}
\afnote{Por ahora uso este $\kket{\rho}$}

We now turn to find the matrix form of the map $\mathcal{E}$ in order to construct the Choi matrix and the conditions the set of parameters $\{\tau_{\alpha}\}$ need to satisfy in order to $\mathcal{E}$ be considered a quantum channel. By writing the density matrix in vector form $\kket{\rho}=2^{-1}\sum_{\alpha=0}^{3} r_{\alpha} \kket{\sigma_{\alpha}}$, the matrix form of the map $\mathcal{E}$ may be written as
% 
\begin{equation}
\hat{\mathcal{E}} = \frac{1}{2}\sum_{\alpha=0}^{3} \tau_{\alpha} \ddyada{\sigma_{\alpha}},
% \label{Map_Pauli}
\end{equation}
%
where the kets $\kket{\sigma_{\alpha}}$ correspond to the vectorized elements of the Pauli basis and satisfy the usual relation $ \bbrakket{{\sigma_{\alpha} }}{{\sigma_{\alpha'} }} = \tr \left(\sigma_{\alpha}^{\dagger} \sigma_{\alpha'}\right) = 2 ~\delta_{\alpha\alpha'}$. After some steps it is possible to show that the associated Choi matrix reads
% 
\begin{equation}
 \mathcal{D} = \frac{1}{2}\sum_{\alpha=0}^{3} \tau_{\alpha} \sigma_{\alpha} \otimes \sigma_{\alpha}^*.
%  \label{Choi_1}
\end{equation}
% 
Moreover, observe that $\mathcal{D}$ has the same form as the matrix representation of a random unitary channel, with Kraus operators given by $K_{\alpha}=\sqrt{\tau_{\alpha}/2}\sigma_{\alpha}$ \cite{CHRUSCINSKI20131425}. In this way the Choi matrix $\mathcal{D}$ is diagonal in the Pauli basis (see Appendix for details),
%
\begin{equation}
 \mathcal{D} \kket{\sigma_{\alpha}}= \lambda_{\alpha} \kket{\sigma_{\alpha}},
\end{equation}
% 
with eigenvalues given by 
% 
\begin{equation}
 \lambda_{\alpha} = \sum_{\beta=0}^3\sa_{\alpha\beta}\tau_{\beta},
\end{equation}
%
or more compactly $\vec{\lambda} = \sa~\vec{\tau}$, where $\sa=2^{-1}H\otimes
H$ and $H$ is the Hadamard matrix,
\cpnote{Decidir si se usa esa $a$ o la otra para ser consistente con lo de los 
generadores}
%
\begin{equation}
 H=\begin{pmatrix}
        1 & 1\\
        1 & -1\\
    \end{pmatrix}.
 \label{Hadd_Mat}
\end{equation}

\cpnote{Mencionar que con la otra $a$ tambien se puede diagonalizar y la
dejo por aca por si se nos ofrece}
\begin{equation}
\sa=\left(
\begin{array}{cccc}
1 &1 &1   &1   \\
 1 & 1  & -1&-1   \\
1  & -1  & 1 & -1\\
1&-1&-1&1  
\end{array}
\right)=\frac12\sa^{-1}.
\label{eq:1}
\end{equation}
% 
%  }}}
\subsubsection{$N$ qubits} % {{{
% 
Analogously to the single qubit case, one can define the $N$-qubits Pauli map $\mathcal{E}_N$
% 
\begin{equation}
\rho \to \rho'=\mathcal{E}_N[\rho ]=\frac{1}{2^N}\sum_{\vec{\alpha}} \tau_{\vec{\alpha}} r_{\vec{\alpha}} \sigma_{\vec{\alpha}},
% \label{Pauli_Map}
\end{equation}
% 
with $-1\leq \tau_{\vec{\alpha}} \leq 1$. 
% 
In this case the density matrix in vector form holds 
% 
\begin{equation}
 \kket{\rho}=\frac{1}{2^N}\sum_{\vec{\alpha}} r_{\vec{\alpha}} \kket{\sigma_{\vec{\alpha} }},
\end{equation}
% 
where $\kket{\sigma_{\vec{\alpha} }}=\kket{\sigma_{\alpha_1}}\otimes\dots
\otimes\kket{\sigma_{\alpha_N}}$, and satisfy the usual orthogonality relation
$\bbrakket{{\sigma_{\vec{\alpha} } }}{{\sigma_{\vec{\alpha}'} }} = \tr
\left(\sigma_{\vec{\alpha}}^{\dagger} \sigma_{\vec{\alpha}'}\right) = 2^N
~\delta_{\vec{\alpha}\vec{\alpha}'}$. The matrix representation of this map is
given by
% 
\begin{equation}
\hat{\mathcal{E}}_N = \frac{1}{2^N}\sum_{\vec{\alpha}} \tau_{\vec{\alpha}} \ddyada{\sigma_{\vec{\alpha} }}.
% \label{Map_Pauli}
\end{equation}
%
As in the previous case, the Choi matrix $\mathcal{D}_N$ may be written in terms of tensor products of Pauli matrices
% 
\begin{equation}
 \mathcal{D}_N = \frac{1}{2^N}\sum_{\vec{\alpha}} \tau_{\vec{\alpha}}  \bigotimes_{j=1}^N \sigma_{\alpha_j} \otimes \sigma_{\alpha_j}^*.
%  \label{Choi_1}
\end{equation}
% 
After several steps, it is possible to prove that analogously to the single qubit case, $\mathcal{D}_N$ is diagonal in the $N$-qubits Pauli basis
%
\begin{equation}
 \mathcal{D}_N \kket{\sigma_{\vec{\alpha} }}
   = \lambda_{\vec{\alpha}} \kket{\sigma_{\vec{\alpha} }},
\end{equation}
% 
with eigenvalues $\lambda_{\vec{\alpha}}$ given by 
% 
\begin{equation}
 \lambda_{\vec{\alpha}} = \sum_{\vec{\beta}} \san_{\vec{\alpha}\vec{\beta}}\tau_{\vec{\beta}},
\end{equation}
%
where 
% 
\begin{equation}
 \san=\sa^{\otimes N}
% =\frac{1}{2^N}\left(H\otimes H\right)^{\otimes N}=\left(\frac{H}{\sqrt{2}}\right)^{\otimes 2N}.
\end{equation}
Again, the proofs are provided in appendix \ref{sec:appendix:diagonalization}
\cpnote{Mencionar que con la otra $a$ tambien se puede diagonalizar para $N$ 
qubits.}
% (Proofs in an appendix using binary indices XD)
% }}}

### 2021-09-29 documentos/vector_space.tex
For the statistics only

\subsection{Parte de Francois} % {{{

We have the matrix 
\begin{equation}
\sa=\left(
\begin{array}{cccc}
1 &1 &1   &1   \\
 1 & 1  & -1&-1   \\
1  & -1  & 1 & -1\\
1&-1&-1&1  
\end{array}
\right)=\frac12\sa^{-1}.
\label{eq:1}
\end{equation}
and further define the matrix
\begin{equation}
\san=\left(\sa\right)^{\otimes N}
\label{eq:1a}
\end{equation}
% }}}
\subsection{A vector space of channels} % {{{
% \begin{itemize} % {{{ Cosas por escribir aca
% \item Canales PCE como espacios vectoriales \todo{Francois escribe}
% \item Regla de $2^k$
% \item Tamaño de las diferentes clases que borran un numero fijo de componentes
% y tambien nota de qeu son igual entre complementarios: computo del numero de
% bases dada una dimension de subespacios vectoriales, aqui entra la formula del
% binomial-q. \todo{Esperemos que evolucione mas el articulo y luego vemos que ponemos}
% \item Intento de identificar elementos, aka hermanistos PCE \cpnote{Esto dependerá de si sale algo interesante,}: De entrada aqui se tiene que mencionar minimo la identificacion del tamaño de las clases.
% \item \ddnote{Agrego esto:} A final remarks of the derivation as a tool.
% Concepts might be complicated, but the tool gives some simple rules to the
% reader: \textit{tell me what components you want to keep, and the (vector)
% summation rules will give you the rest of components to make the target
% operation a CPTP one.}
% \end{itemize} % }}}

% Inversion de la relacion entre tau y p, y que las probablidiades deben ser positivas {{{
The problem of determining complete positivity of a PCE can be recast as 
determining which vectors $\vec \tau$ are mapped via
$\san$
% \begin{equation}
% \san\vec\tau=2^N\vec p
% \label{eq:condition}
% \end{equation}
to a positive vector $\vec p$. In compontents, this would read 
\begin{equation}
% \left(
\san_{\vec\alpha\vec\beta} \,
% \right)
\tau_\beta=2^Np_{\vec\alpha}
\label{eq:A:tau:is:p}
\end{equation}
where
$\san_{\vec\alpha\vec\beta} =\sa_{\alpha_1\beta_1}\cdot\ldots\cdot
\sa_{\alpha_N\beta_N}$. Using the fact that $a=a^{-1/2}$ (so
$A = A^{-1}/2^N$) we can directly invert \eref{eq:A:tau:is:p} 
and obtain 
\begin{equation}
\tau_\alpha=
% \left(
\san_{\vec\alpha\vec\beta} \,
% \right)
p_\beta,
\label{eq:tau:is:A:p}
\end{equation}
which will serve as a starting point for our analysis. Other simple but crucial
observation that we shall use is that, for valid quantum channels, 
$p_{\vec\alpha}+p_{\vec\beta}=0$ implies $p_{\vec\alpha}=p_{\vec\beta}=0$, as
all $p_{\vec\alpha} \ge 0$ due to 
complete positivity of the underlying channel.

% We shall essentially use the fact that
% 
% 
% 
% It results convenient to express the 
% previous equation in  
%

% We look for those vectors $\vec\tau$ consisting of 0's and 1's such that 
% \begin{equation}
% \san\vec\tau=2^N\vec p
% \label{eq:3}
% \end{equation}
% where $\vec p$ is a vector with positive entries. Here $\vec\tau$ is a $4^N$ dimensional vector of 0's
% and 1's. We number the entries of $\vec\tau$ 
% by the multi-indices $\vec\alpha=(\alpha_1,\ldots,\alpha_N)$ where $0\leq\alpha_k\leq3$.
% The entries of the positive vector are given by $2^Np_{\vec\alpha}$. It then follows that 
% the expression (\ref{eq:3}) satisfies
% \begin{equation}
% \left(
% \san_{\vec\alpha\vec\beta}
% \right)
% \tau_\beta=2^Np_{\vec\alpha}
% \label{eq:4}
% \end{equation}
% where
% \begin{equation}
% \left(
% \san_{\vec\alpha\vec\beta}
% \right)=\sa_{\alpha_1\beta_1}\cdot\ldots\cdot \sa_{\alpha_N\beta_N}
% \label{eq:5}
% \end{equation}
% 
% From (\ref{eq:4}) immediately follows, via (\ref{eq:1})
% \begin{equation}
% \tau_\alpha=\left(
% \san_{\vec\alpha\vec\beta}
% \right)p_\beta
% \label{eq:6}
% \end{equation}
% which we analyze iteratively. We shall essentially use the fact that
% \begin{equation}
% p_{\vec\alpha}+p_{\vec\beta}=0\quad\Leftrightarrow\quad p_{\vec\alpha}=p_{\vec\beta}=0
% \label{eq:7}
% \end{equation}
% }}}

Now we need a definition: to each multi-index $\vec\alpha$ we associate a {\em set\/}
of multi-indices $\Phi(\vec\alpha)$ as follows
\begin{equation}
\Phi(\vec\alpha):=\left\{
\vec\beta:\san_{\vec\alpha\vec\beta}=1
\right\}=\left\{
\vec\beta:
\sa_{\alpha_1\beta_1}\cdot\ldots\cdot \sa_{\alpha_N\beta_N}
=1
\right\}
\label{eq:8}
\end{equation}
If we now assume that $\tau_{\vec\alpha}=1$, it follows from subtracting the two combined equations
\begin{subequations}
\begin{eqnarray}
\sum_{\vec\beta}p_{\vec\beta}&=&1
\label{eq:9a}\\
\sum_{\vec\beta}\san_{\vec\alpha\vec\beta}p_{\vec\beta}&=&\tau_{\vec\alpha}=1
\end{eqnarray}
\label{eq:9}
\end{subequations}
that for all $\vec\beta\notin\Phi(\vec\alpha)$
\begin{eqnarray}
p_{\vec\beta}=0.
\label{eq:10}
\end{eqnarray}
Thus, if $\tau_{\vec\alpha}=1$, then $\tau_{\vec\gamma}$
and $\tau_{\vec{\gamma^\prime}}$ are equal if, for all $\vec\beta\in\Phi(\vec\alpha)$.
\begin{equation}
\san_{\vec\beta\vec\gamma}=\san_{\vec\beta\vec{\gamma^\prime}}
\label{eq:11}
\end{equation}
holds for all $\vec\beta\in\Phi(\vec\alpha)$. This follows from (\ref{eq:6}) as well as the fact that, due 
to (\ref{eq:10}), we only need sum over $\vec\beta\in\Phi(\vec\alpha)$.

>>>>

\subsection{A vector space of channels} % {{{
% \begin{itemize} % {{{ Cosas por escribir aca
% \item Canales PCE como espacios vectoriales \todo{Francois escribe}
% \item Regla de $2^k$
% \item Tamaño de las diferentes clases que borran un numero fijo de componentes
% y tambien nota de qeu son igual entre complementarios: computo del numero de
% bases dada una dimension de subespacios vectoriales, aqui entra la formula del
% binomial-q. \todo{Esperemos que evolucione mas el articulo y luego vemos que ponemos}
% \item Intento de identificar elementos, aka hermanistos PCE \cpnote{Esto dependerá de si sale algo interesante,}: De entrada aqui se tiene que mencionar minimo la identificacion del tamaño de las clases.
% \item \ddnote{Agrego esto:} A final remarks of the derivation as a tool.
% Concepts might be complicated, but the tool gives some simple rules to the
% reader: \textit{tell me what components you want to keep, and the (vector)
% summation rules will give you the rest of components to make the target
% operation a CPTP one.}
% \end{itemize} % }}}

% Inversion de la relacion entre tau y p, y que las probablidiades deben ser positivas {{{
The problem of determining complete positivity of a PCE can be recast as 
determining which vectors $\vec \tau$ are mapped via
$\san$
% \begin{equation}
% \san\vec\tau=2^N\vec p
% \label{eq:condition}
% \end{equation}
to a positive vector $\vec p$. In compontents, this would read 
\begin{equation}
% \left(
\san_{\vec\alpha\vec\beta} \,
% \right)
\tau_\beta=2^Np_{\vec\alpha}
\label{eq:A:tau:is:p}
\end{equation}
where
$\san_{\vec\alpha\vec\beta} =\sa_{\alpha_1\beta_1}\cdot\ldots\cdot
\sa_{\alpha_N\beta_N}$. Using the fact that $a=a^{-1/2}$ \cpnote{ojo cno factores
globales, depende de la definicion de la $a$ y $A$} (so
$A = A^{-1}/2^N$) we can directly invert \eref{eq:A:tau:is:p} 
and obtain 
\begin{equation}
\tau_\alpha=
% \left(
\san_{\vec\alpha\vec\beta} \,
% \right)
p_\beta,
\label{eq:tau:is:A:p}
\end{equation}
which will serve as a starting point for our analysis. Other simple but crucial
observation that we shall use is that, for valid quantum channels, 
$p_{\vec\alpha}+p_{\vec\beta}=0$ implies $p_{\vec\alpha}=p_{\vec\beta}=0$, as
all $p_{\vec\alpha} \ge 0$ due to 
complete positivity of the underlying channel. 
% }}}

Now we need a definition: to each multi-index $\vec\alpha$ we associate a \textit{set}
of multi-indices $\Phi(\vec\alpha)$ as follows
\begin{equation}
\Phi(\vec\alpha):=\left\{
\vec\beta:\san_{\vec\alpha\vec\beta}=1
\right\}
% =\left\{ \vec\beta: \sa_{\alpha_1\beta_1}\cdot\ldots\cdot \sa_{\alpha_N\beta_N}
% =1
% \right\}
\label{eq:8}
\end{equation}
If we now assume that $\tau_{\vec\alpha}=1$, it follows from subtracting the two combined equations
\begin{subequations}
\begin{eqnarray}
\sum_{\vec\beta}p_{\vec\beta}&=&1
\label{eq:9a}\\
\sum_{\vec\beta}\san_{\vec\alpha\vec\beta}p_{\vec\beta}&=&\tau_{\vec\alpha}=1
\end{eqnarray}
\label{eq:9}
\end{subequations}
that for all $\vec\beta\notin\Phi(\vec\alpha)$
\begin{eqnarray}
p_{\vec\beta}=0.
\label{eq:10}
\end{eqnarray}
Thus, if $\tau_{\vec\alpha}=1$, then $\tau_{\vec\gamma}$
and $\tau_{\vec{\gamma^\prime}}$ are equal if, for all $\vec\beta\in\Phi(\vec\alpha)$.
\begin{equation}
\san_{\vec\beta\vec\gamma}=\san_{\vec\beta\vec{\gamma^\prime}}
\label{eq:11}
\end{equation}
holds for all $\vec\beta\in\Phi(\vec\alpha)$. This follows from (\ref{eq:6}) as well as the fact that, due 
to (\ref{eq:10}), we only need sum over $\vec\beta\in\Phi(\vec\alpha)$.
